---
title: "Homework 3 - Visualization and EDA"
author: "Sydney Ng (uni: sn2863)"
date: "due 10/10/2020 by 10:00 pm EDT"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)
```

## Problem 1 - Instacart data
```{r}
data("instacart") # loading in the data set
```

This data set contains `r nrow(instacart)` number of observations and `r ncol(instacart)` variables.

Observations are the level of items in orders by users. In the data set, there are user and order variables, including user ID, order ID, order day, and order hour. Other variables are aisles with corresponding aisle ID, which belong to product departments.

#### How many aisles and which are the most from?
```{r}
aisle_distinct <- 
  instacart %>%
  group_by(aisle) %>%
  summarize(count = n()) %>%
  arrange(desc(count))

nrow(aisle_distinct) # there are 134 different aisles
```

#### Let's make a plot!
```{r}
aisle_distinct %>%
  filter(count > 10000) %>%
  mutate(aisle = factor(aisle),
         aisle = fct_reorder(aisle,count)) %>%
  ggplot(aes(x = aisle, y = count)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 270, vjust=0.5, hjust=0))
```

Table showing the three most popular items in each of the aisles "baking ingredients", "dog food care", and "packaged vegetables fruits". Include the number of times each item is ordered in your table.

```{r}
popular_items <-
  instacart %>%
  filter(aisle %in% c("baking ingredients", 
                      "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank)
```

Another table for Pink Lady Apples versus Coffee Ice Cream
```{r}
apples_icecream <-
  instacart %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(names_from = order_dow, values_from = mean_hour)
  # deliberately untidy the data set -- long to wide format

```



## Problem 2
```{r}
accel_df <- 
  read_csv("./accel_data.csv") %>%
  pivot_longer(4:1443, names_to = "activity", values_to = "activity_count",
               names_prefix = "activity.") %>%
  mutate(weekend = if_else(day == "Saturday" |
                             day == "Sunday", 1, 0), # weekday vs weekend
         activity = as.numeric(activity)) #changing chr to numeric class
```

After cleaning the `accel_data.csv` we have `r nrow(accel_df)` observations in the data set and `r ncol(accel_df)` variables.


## Problem 3
